Simulation connectance richness

### 2023-07-06 report/cs_analysis.Rmd
Improve intro

Understanding the drivers of temporal stability is a key question in Ecology.
BEF theory shows that species richness is a key driver to understand the
temporal stability of community biomass.

Species richness was demonstrated to increase community stability by driving higher
asynchrony in community in various assemblages and ecosystems. In particular,
high response diversity, which drives independant species fluctuations, seems to a key
factor driving the positive effect of species richness of stability
[@loreau_species_2008; @thibaut_understanding_2013; @zhao_biodiversity_2022]. However,
those studies were often limited to monotrophic assemblages. As natural
communities are often composed of multiple trophic level, our understanding of
community stability needs to incorporate food-web structure.

Food-web structure has been shown to have tremendous effect on community
stability. The degree of generalism of predator was shown to modulate
stability-richness relationships [@thebault_trophic_2005]. The strength and distribution of
interaction strength was also shown to be strong driver of community stability
[@mccann_weak_1998]. Omnivory was also shown to drive higher stability
[@mccann_reevaluating_1997]. However, it is quite unkown how those food-web
structure elements would affect stability-species richness relationships.

The effect of environmental stressors can also strongly affect community
variability. Higher environmental stress and variability may dampens the effect
of richness on stability [@loreau_biodiversity_2001]. Lastly, productivity of a
system, as the quantity of biomass input in the ecosystem, can have tremendous
effect on both community structure and stabilty [@lindeman_trophic-dynamic_1942;
@binzer_dynamics_2012; @binzer_interactive_2016]. Productivity can increase
species richness in the food-web by increasing the level of energy input
[@lindeman_trophic-dynamic_1942] but also destabilises community biomass by
increasing population variability, a behavior coined as the paradox of
enrichment [@rosenzweig_paradox_1971]. A better understanding of the effect of
species richness on community stability should then include environmental
gradient and stressors.

Community stability can itself be partitioned in several components, i.e.
population stability and population asynchrony [@thibaut_understanding_2013;
@zhao_biodiversity_2022]. Population stability can be itself partioned in total
biomass and population variability effect. Population asynchrony can further be
partioned in statistical averaging effect and compensatory dynamics effects
[@zhao_biodiversity_2022]. Partitioning stability allows to better understand
how food-web structure and external factors affect stability. For exemple,
environmental stochasticity is expected to primarily affect population
stability, while response diversity should affect primarily compensatory
dynamics.

@@

```{r}
tribble(
  ~Type, ~`Food-web structure`, ~`Effect on stability-richness`,
  "Food-web structure", "Connectance", "Strengthen (Miele et al., 2016) / No effect (Thébault & Loreau, 2005)",
  "Food-web structure", "Omnivory", "Strengthen (Mccann et al., 1997)",
  "Food-web structure", "Interaction strength", "Destabilising (Mccann et al., 1998) / Stabilising in
  productive (Binzer et al., 2012; Binzer et al., 2016; Tabi et al., 2019)",
  "Response diversity", "", "Dampens",
  "Productivity", "", "Dampens (Binzer et al., 2012; Binzer et al., 2016; Tabi et al., 2019)",
  "Environmental variability", "", "Dampens",
)
```

@@

```{r}
sim_ok %>%
  filter(ct_alive != 0, rho == 0, env_stoch == .1)
sim_sem <- sim_ok %>%
  filter(ct_alive != 0) %>%
  mutate(across(c(stab_com, async, pop_stab, total_biomass, inv_sum_pop_var, sae_even, evenness_sae, cpe_env, cpe_int), log, , .names = "log_{.col}")) %>%
  na.omit()
```

@@

```{r}
sim_stab %>%
  ggplot(aes(x = as.factor(rho), y = cpe_env)) +
  geom_boxplot() +
  geom_hline(yintercept = 1)
sim_stab %>%
  ggplot(aes(x = as.factor(env_stoch), y = cpe)) +
  geom_boxplot() +
  geom_hline(yintercept = 1)

sim_stab %>%
  ggplot(aes(x = as.factor(env_stoch), y = stab_com)) +
  geom_boxplot() +
  geom_hline(yintercept = 1)
sim_stab %>%
  ggplot(aes(x = as.factor(rho), y = stab_com)) +
  geom_boxplot() +
  geom_hline(yintercept = 1)
sim_stab %>%
  ggplot(aes(x = as.factor(rho), y = evenness_sae)) +
  geom_boxplot() +
  geom_hline(yintercept = 1)
sim_stab %>%
  ggplot(aes(x = as.factor(rho), y = sae_total)) +
  geom_boxplot() +
  geom_hline(yintercept = 1)

>>>>

Global change and their effects constitute major treats to ecosystem stability.
Among them, global change is predicted to lead dramatically more variable
environments [@pachauri_climate_2015]. This new threat adds to current
eutrophication in many ecosystems [@maure_globally_2021, add ref], which is
known to dramatically impair ecosystem functioning. Understanding how
environmental variability and eutrophication will alter ecosystem stability is
then a timely question.

Biodiversity, in terms of species richness, has been repeatedly shown to drive
higher temporal stability of ecological communities, across various taxa and
ecosystems. Species richness can dampen the effects of perturbations on the
temporal variability of community biomass because more diverse communities
contain species with more diversity of responses to perturbations, which
generates asynchrony in population fluctuations, thereby stabilizing community
biomass fluctuations [@loreau_species_2008; @thibaut_understanding_2013;
@zhao_biodiversity_2022]. In turn, the effects of human pressures on biotic
homogenization reported in various ecosystems [@buisson_toward_2013;
@muthukrishnan_invasive_2020; @saladin_rapid_2020] is another threat, which
might in turn affect dampens the effects of species richness on ecosystem
stability. However, the joint effects of external stressors such as
environmental variability and eutrophication and internal stressors such as the
loss response diversity has been rarely assessed in temporal stability areas.

The question of stressors can even become complex in multitrophic communities,
where species richness corresponds both to horizontal and vertical diversity.
Food-web structure has been shown to have tremendous effect on community
stability. The degree of generalism of predator was shown to modulate
stability-richness relationships [@thebault_trophic_2005]. The strength and
distribution of interaction strength was also shown to be strong driver of
community stability [@mccann_weak_1998]. Omnivory was also shown to drive higher
stability [@mccann_reevaluating_1997]. The effects of food-web structure on
ecosystem stability has been mostly assessed with small modules [but see
@eschenbrenner_diversity_nodate], and we know very little how food-web structure
affects the relationships between species richness and temporal stability.

Both stressors and food-web structure are expected to have different effects on
temporal stability. Stability of community biomass can itself be partitioned in
several components, i.e. population stability and population asynchrony
[@thibaut_understanding_2013; @zhao_biodiversity_2022]. Population stability can
be itself partioned in total biomass and population variability effect.
Population asynchrony can further be partioned in statistical averaging effect
and compensatory dynamics effects [@zhao_biodiversity_2022]. Partitioning
stability allows to better understand how food-web structure and external
factors affect stability. We expect that environmental variability will
primarily affect. For exemple, environmental stochasticity is expected to
primarily affect population variance. Productivity is expected to have two
effects on stability. First, productivity may increase species richness and
average trophic level by increasing the flux of biomass in the system
[@lindeman_trophic-dynamic_1942] but on the other side, the resulting higher growth rate
of primary producers may increase the population variability
[@binzer_interactive_2016] and thus decreases the temporal stability of
community biomass. Response diversity is expected to act primarily on
compensatory dynamics in the food-webs, as low response diversity will decrease
the effects. Higher connectance and interaction strength are expected to dampens
population variability but to lead to higher synchrony at the same time.

---
#Understanding the mechanisms that determine temporal
#stability of ecosystem functioning is a long standing
#question in Ecology [@may_will_1972; @angelis_stability_1975;
#@mccann_diversitystability_2000; @tilman_biodiversity_2006;
#@donohue_navigating_2016; @kefi_advancing_2019].
---

In this study, we investigated both the effects of stressors (productivity,
environnmental variability and response diversity) and of food-web structure
(connectance, average interaction strength, average trophic level) on species
richness-stability relationships. To do, we used a bioenergetic food-web model
[@yodzis_body_1992; @brose_allometric_2006; @delmas_simulations_2017] added to
environmental stochasticity.

@@

```{r}
tribble(
  ~Type, ~`Food-web structure`, ~`Effect on stability-richness`,
  "Food-web structure", "Connectance", "Strengthen (Miele et al., 2016) / No effect (Thébault & Loreau, 2005)",
  "Food-web structure", "Omnivory", "Strengthen (Mccann et al., 1997)",
  "Food-web structure", "Interaction strength", "Destabilising (Mccann et al., 1998) / Stabilising in
  productive (Binzer et al., 2012; Binzer et al., 2016; Tabi et al., 2019)",
  "Response diversity", "", "Dampens",
  "Productivity", "", "Dampens (Binzer et al., 2012; Binzer et al., 2016; Tabi et al., 2019)",
  "Environmental variability", "", "Dampens",) %>%
kbl(caption = "Predictions of stability-richness relationship from the literature.") %>%
kable_styling()
```

@@

```{r}
sim_sem <- sim_ok %>%
  filter(max_tlvl_alive > 1) %>%
  mutate(across(c(stab_com, async, pop_stab, total_biomass, inv_sum_pop_var, sae_even, evenness_sae, cpe_env, cpe_int), log, , .names = "log_{.col}")) %>%
  na.omit()
```

```{r}
summary(sim_sem$max_tlvl_alive)
```

```{r}
p_sim_rich_prod_fw_only <- sim_sem %>%
  mutate(stab_pop = 1/ avg_cv_sp) %>%
  filter(rho %in% c(0, 1)) %>%
  pivot_longer(
    c(stab_com, stab_pop, sync),
    names_to = "metric",
    values_to = "stab") %>%
  mutate(
    metric = var_replacement()[metric],
    rho = rho_replacement()[as.character(rho)]
  ) %>%
  ggplot(aes(x = log(richness), y = log(stab), color = factor(productivity))) +
  geom_point(alpha = .1, shape = 20) +
  geom_smooth(method = "lm") +
  facet_grid(rows = vars(metric), cols = vars(rho), scales = "free_y") +
  labs(
    color = "Productivity",
    x = "Species richness",
    y = "Stability component"
  ) +
  scale_x_continuous(
    sec.axis = sec_axis(~ . ,
      name = "Response diversity",
      breaks = NULL,
      labels = NULL)
  )
ggsave(
  filename = here("report", "figures", "p_sim_rich_prod_fw_only.png"),
  p_sim_rich_prod_fw_only,
  scale = 1.4,
  width = 88,
  height = 88 * 1,
  units = "mm"
)
colnames(sim_sem)
```

```{r}
p_sim_rich_prod_fw_only_cpe <- sim_sem %>%
  mutate(stab_pop = 1/ avg_cv_sp, ) %>%
  filter(rho %in% c(0, 1)) %>%
  pivot_longer(
    c(sae_total, cpe),
    names_to = "metric",
    values_to = "stab") %>%
  mutate(
    metric = var_replacement()[metric],
    rho = rho_replacement()[as.character(rho)]
  ) %>%
  ggplot(aes(x = log(richness), y = log(stab), color = factor(productivity))) +
  geom_point(alpha = .1, shape = 20) +
  geom_smooth(method = "lm") +
  facet_grid(rows = vars(metric), cols = vars(rho), scales = "free_y") +
  labs(
    color = "Productivity",
    x = "Species richness",
    y = "Stability component"
  ) +
  scale_x_continuous(
    sec.axis = sec_axis(~ . ,
      name = "Response diversity",
      breaks = NULL,
      labels = NULL)
  )
ggsave(
  filename = here("report", "figures", "p_sim_rich_prod_fw_only_cpe.png"),
  p_sim_rich_prod_fw_only_cpe,
  scale = 1.4,
  width = 88,
  height = 88 * 1,
  units = "mm"
)
```

@@

```{r}
sim_ok %>%
  filter(max_tlvl_alive > 1)
```

```{r}
sim_stab %>%
  ggplot(aes(x = as.factor(rho), y = cpe_env)) +
  geom_boxplot() +
  geom_hline(yintercept = c(.5, 1))
sim_stab %>%
  ggplot(aes(x = as.factor(env_stoch), y = cpe_env)) +
  geom_boxplot() +
  geom_hline(yintercept = c(.5, 1))
sim_stab %>%
  ggplot(aes(x = as.factor(env_stoch), y = cpe)) +
  geom_boxplot() +
  geom_hline(yintercept = 1)

sim_stab %>%
  ggplot(aes(x = as.factor(env_stoch), y = stab_com)) +
  geom_boxplot() +
  geom_hline(yintercept = 1)
sim_stab %>%
  ggplot(aes(x = as.factor(rho), y = stab_com)) +
  geom_boxplot() +
  geom_hline(yintercept = 1)
sim_stab %>%
  ggplot(aes(x = as.factor(rho), y = evenness_sae)) +
  geom_boxplot() +
  geom_hline(yintercept = 1)
sim_ok %>%
  filter(max_tlvl_alive > 1) %>%
  ggplot(aes(x = as.factor(as.integer(richness)), y = evenness_sae)) +
  geom_boxplot() +
  geom_hline(yintercept = 1)
sim_ok %>%
  filter(max_tlvl_alive > 1) %>%
  ggplot(aes(x = as.factor(as.integer(max_tlvl_alive)), y = evenness_sae)) +
  geom_boxplot() +
  geom_hline(yintercept = 1)
sim_stab %>%
  ggplot(aes(x = as.factor(rho), y = sae_total)) +
  geom_boxplot() +
  geom_hline(yintercept = 1)
